Model-based Clustering for Neural Populations

### 2021-09-27 documents/cluster_neural_population.tex
documentation

In summary, Let
\(\mathbf{\Theta} = \left\{ \mathbf{z},\ \mathbf{d},\mathbf{C},\left\{ \mathbf{x}_{t} \right\}_{t = 1}^{T},\mathbf{x}_{0},\mathbf{Q}_{0},\mathbf{A},\ \mathbf{b},\mathbf{Q}\  \right\}\)
be the set of parameters for LDS. The number of observation/ neuron is
\(N\), and they can group into \(J\) clusters. In each cluster, there
are \(p\) latent state vectors. The recording length is \(T\).
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	\(\mathbf{z} \in \mathbb{Z}^{N}\): cluster index for each neuron/ observation,
	with \(z_{i} \in \{ 1,\ldots,J\}\) for \(i = 1,\ldots,N\).
	\item
	\(\mathbf{d} \in \mathbb{R}^{N}\) and \(\mathbf{C} \in \mathbb{R}^{N \times p}\): baseline
	\& loading of the latent
	\item
	\(\mathbf{x}_{t} \in \mathbb{R}^{\text{Jp}}\): all latent at \(t\).
	\item
	\(\mathbf{x}_{t}^{(j)} \in \mathbb{R}^{p}\) means the latent in cluster \(j\) at
	time \(t\).
	\item
	\(\mathbf{x}_{0} \in \mathbb{R}^{\text{Jp}}\) and
	\(\mathbf{Q}_{0} \in \mathbb{R}^{\text{Jp} \times Jp}\): mean and covariance of
	\(\mathbf{x}_{1}\). For simplicity, assume \(\mathbf{Q}_{0}\) is known.
	\item
	\(\mathbf{A} \in \mathbb{R}^{Jp \times Jp},\ \mathbf{b} \in \mathbb{R}^{Jp},\mathbf{Q} \in \mathbb{R}^{Jp \times Jp}\):
	linear dynamics of the latent.
\end{enumerate}

@@

The conditional priors for these parameters:
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	Latent vectors \(\left\{ \mathbf{x}_{t} \right\}_{t=1}^T\): the conditional prior is
	defined by
	\[\mathbf{x}_{1} \sim N(\mathbf{x}_{0},\mathbf{Q}_{0})\]
	\[\mathbf{x}_{t + 1}|\mathbf{x}_{t} \sim N(\mathbf{A}\mathbf{x}_{t} + \mathbf{b},\mathbf{Q})\]
	\item
	Initials \(\mathbf{x}_{0}\): assume there are \(J\) clusters,
	\[\mathbf{x}_{0} \sim N(\bm{\mu}_{\mathbf{x}_{00}},\ \mathbf{\Sigma}_{\mathbf{x}_{00}})\]
	, where \(\bm{\mu}_{\mathbf{x}_{00}} = \mathbf{0}_{\text{Jp}}\) and \(\mathbf{\Sigma}_{\mathbf{x}_{00}} = \mathbf{I}_{\text{Jp}}\)
	\item
	Linear mapping (loading) for latent vectors
	\(\left\{ d_{i} \right\}_{i = 1}^{N}\) and
	\(\left\{ \mathbf{c}_{i} \right\}_{i = 1}^{N}\):
	\[\left( d_{i},\mathbf{c}'_{i} \right)' \sim N(\bm{\mu}_{\text{dc}}^{\left( z_{i} \right)},\mathbf{\Sigma}_{\text{dc}}^{(z_{i})})\]
	\item
	Mean and covariance for loading in each cluster
	Mean and covariance for loading in each cluster
	\(\left\{ \bm{\mu}_{\text{dc}}^{(j)} \right\}_{j}\) and
	\(\left\{ \mathbf{\Sigma}_{\text{dc}}^{(j)} \right\}_{j}\):
	\[\bm{\mu}_{\text{dc}}^{(j)} \sim N(\mathbf{\delta}_{dc0},\mathbf{T}_{dc0})\]
	, where \(\bm{\delta}_{dc0} = \mathbf{0}_{p + 1}\) and
	\(\mathbf{T}_{dc0} = \mathbf{I}_{p + 1}\)

@@

	, where \(\mathbf{f}_0 = vec(\mathbf{F}_0)\).If the number of cluster is \(J\), \(\nu_{\mathbf{Q}_{0}} = Jp + 2\),
	\(\Psi_{\mathbf{Q}_{0}} = \mathbf{I}_{\text{Jp}} \times 10^{- 4}\). (To make the mean of \(\mathbf{Q}\) loosely centered around
	\(\mathbf{I}_{Jp} \times 10^{- 4}\)), \(\mathbf{F}_0 = \begin{pmatrix}
		\mathbf{0}'_{Jp} \\
		\mathbf{I}_{\text{Jp}} \\
	\end{pmatrix}\) and \(\bm{\Lambda}_0 = \mathbf{I}_{Jp + 1}\)

@@

, where
\(\mathbf{E} = \left( \mathbf{\epsilon}_{1},\ldots,\mathbf{\epsilon}_{T - 1} \right)^{'}\) and \(\mathbf{\epsilon} \sim N(0,\mathbf{Q})\). Then posteriors are
\begin{align*}
	\mathbf{Q}|\mathbf{Y}_{\mathbf{\text{bA}}},\mathbf{X}_{\mathbf{\text{bA}}},\ldots &\sim W^{-1}(\Psi_{\mathbf{Q}}, \nu_{\mathbf{Q}})\\
	\mathbf{f}|\mathbf{Y}_{\mathbf{\text{bA}}},\mathbf{X}_{\mathbf{\text{bA}}},\mathbf{Q},\ldots &\sim N(vec(\mathbf{F}_\mathbf{bA}),  \mathbf{Q}\bigotimes \bm{\Lambda}_\mathbf{bA}^{-1})
\end{align*}
, with parameters be:
\begin{align*}
	\Psi_{\mathbf{Q}} &= \Psi_{\mathbf{Q}_{0}} + (\mathbf{Y}_\mathbf{bA} - \mathbf{X}_\mathbf{bA}\mathbf{F}_\mathbf{bA})'(\mathbf{Y}_\mathbf{bA} - \mathbf{X}_\mathbf{bA}\mathbf{F}_\mathbf{bA}) + (\mathbf{F}_\mathbf{bA} - \mathbf{F}_0)'\bm{\Lambda}_0(\mathbf{F}_\mathbf{bA} - \mathbf{F}_0)\\
	\nu_{\mathbf{Q}} &= \nu_{\mathbf{Q}_{0}} + T - 1\\
	\mathbf{F}_\mathbf{bA} &= (\mathbf{X}'_\mathbf{bA}\mathbf{X}_\mathbf{bA} + \bm{\Lambda}_0)^{-1}(\mathbf{X}'_\mathbf{bA}\mathbf{Y}_\mathbf{bA} + \bm{\Lambda}_0\mathbf{F}_0)\\
	\bm{\Lambda}_\mathbf{bA} &= \mathbf{X}'_\mathbf{bA}\mathbf{X}_\mathbf{bA} + \bm{\Lambda}_0
\end{align*}

\section{Model Check [Problems for \(\mathbf{b}\) and \(\mathbf{A}\) partially solved]} \label{problems}

\subsection{Estimation of \(\mathbf{d}\) and \(\mathbf{C}\) alone}
First, turn off all others except for loading (related) parameters and set them to be true values. There are 2 versions: (1) update priors as in subsection \ref{loading prior} and (2) no update of priors, and just set the prior for each as the standard normal.
For the estimation with prior update, the results are mean from iteration 50 to 100 (Figure \ref{loading alone}).
\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[]{\includegraphics[width=0.6\textwidth]{image016.png}}%
		\subfigure[]{\includegraphics[width=0.6\textwidth]{image017.png}}%
	}
	\caption{estimation of loading}
	\label{loading alone}
\end{figure}

The mixing of convergence for loading priors (in subsection \ref{loading prior}) are fine. The update of priors don't influence results a lot, but this will help clustering a lot. It seems estimation of loading is fine. Let's see what happens when estimation of latents \(\mathbf{x}_t\) is on at the same time.

\subsection{Estimation of \(\mathbf{d}\), \(\mathbf{C}\) and \(\mathbf{x}_{t}\)}
When the estimation of \(\mathbf{x}_t\) is on, we need more iterations. The results are mean from iteration 500 to 1000 (Figure \ref{loading and latent}).

The estimation of loading:
\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[]{\includegraphics[width=0.6\textwidth]{image018.png}}%
		\subfigure[]{\includegraphics[width=0.6\textwidth]{image020.png}}%
	}
	\caption{estimation of loading, with latents on}
	\label{loading and latent}
\end{figure}

And the estimation of corresponding latents (Figure \ref{loading and latents}):
\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[no update of prior]{\includegraphics[width=0.6\textwidth]{image019.png}}%
		\subfigure[with update of prior]{\includegraphics[width=0.6\textwidth]{image021.png}}%
	}
	\caption{estimation of loading}
	\label{loading and latents}
\end{figure}

Things are still fine. The remaining parts may influence latent estimations are dynamics (\(\mathbf{b}\), \(\mathbf{A}\)) and prior process noise \(\mathbf{Q}\).

\subsection{Estimation of \(\mathbf{b}\), \(\mathbf{A}\), \(\mathbf{Q}\) and \(\mathbf{x}_t\)}
I show the results of block-diagonal (Figure \ref{bAQX_blkDiag}) and full (Figure \ref{bAQX_full}) \(\mathbf{Q}\). These results are from 2000 MCMC samples. The latent and dynamics are from average from 500 to 2000 iterations.

\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[trace]{\label{fig:a}\includegraphics[width=0.55\textwidth]{bAX_blkDiag_trace.png}}%
		\subfigure[latent]{\label{fig:b}\includegraphics[width=0.55\textwidth]{bAX_blkDiag_latent.png}}%
		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{bAX_blkDiag_A.png}}%
	}
	\caption{sampling of $\mathbf{b}$, $\mathbf{A}$, $\mathbf{Q}$ and $\mathbf{x}_t$, block-diagonal process noise}
	\label{bAQX_blkDiag}
\end{figure}

\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[trace]{\label{fig:a}\includegraphics[width=0.55\textwidth]{bAX_full_trace.png}}%
		\subfigure[latent]{\label{fig:b}\includegraphics[width=0.55\textwidth]{bAX_full_latent.png}}%
		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{bAX_full_A.png}}%
	}
	\caption{sampling of $\mathbf{b}$, $\mathbf{A}$, $\mathbf{Q}$ and $\mathbf{x}_t$, full process noise}
	\label{bAQX_full}
\end{figure}

Both block-diagonal and full looks fine, but the mixing of  full process noise version looks better.

@@

There are two set of simulation examples. The first example is generated from LDS model directly, while in the second example the latents are generated directly without explicit specifying linear dynamics. In all following results, I fit things with both \href{https://github.com/weigcdsb/state-space-clustering/tree/main/LDS/blkDiag}{block-diagonal} and \href{https://github.com/weigcdsb/state-space-clustering/tree/main/LDS/full}{full} \(\mathbf{Q}\). The fitting results for block-diagonal version is a bit better, because the underlying true \(\mathbf{Q}\) is diagonal. In the following part, I only show results from block-diagonal fitting for labeled data. For unlabeled data, i.e. clustering, all the results can be found in this \href{https://github.com/weigcdsb/state-space-clustering/tree/main/results/gif}{folder}.

\subsection{Simulation 1: Generate from LDS Directly}
In this simulation, there are 3 clusters with 10 neurons in each cluster. The dimension of latents in each cluster is 2. In the linear dynamics, the bias term \(\mathbf{b}\) is zero. There's no within-population interaction but has some weak between-population interactions. In other words, the linear dynamics matrix \(\mathbf{A}\) is roughly diagonal. The details of simulation can be found in the first section of the \href{https://github.com/weigcdsb/state-space-clustering/blob/main/LDS/blkDiag/lds_sample_DP_blkDiag.m}{simulation 1}.

\subsubsection{Labeled Data: No Clustering}
The code for fitting is \href{https://github.com/weigcdsb/state-space-clustering/blob/main/LDS/blkDiag/lds_sample_DP_blkDiag.m}{here}.

Here, I first compare 10k iterations for three methods (consider high dimension of \(\mathbf{x}_t\)): (1) both \(\mathbf{x}_t\) and loading (\(\mathbf{d}\) and \(\mathbf{C}\)) are updated by normal approximation. (2) update \(\mathbf{x}_t\) by normal approximation, but update loading by MH. (3) update both \(\mathbf{x}_t\) and loading by MH.

The fitting results are average of iteration 5k to 10k.

\textbf{Update \(\mathbf{x}_t\) and (\(\mathbf{d}\), \(\mathbf{C}\)) by normal approximation}

The trace plots are in Figure \ref{double N: trace}
\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[tracd of latent]{\label{fig:a}\includegraphics[width=0.55\textwidth]{xnorm_doubleN.png}}%
		\subfigure[trace of loading and dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{trace_doubleN.png}}%
	}
	\caption{both Normal Approx., trace plots}
	\label{double N: trace}
\end{figure}

Well, the chain hasn't achieve the stationary distribution. So there's no need to see further results. But to show the improvement and for comparison, I still show the fitting of overall firing rate, latents and dynamics in Figure \ref{double N: fit}.

\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[overall mean firing rate]{\label{fig:a}\includegraphics[width=0.55\textwidth]{FR_doubleN.png}}%
		\subfigure[latents]{\label{fig:b}\includegraphics[width=0.55\textwidth]{latent_doubleN.png}}%
		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{A_doubleN.png}}%
	}
	\caption{both Normal Approx., fit}
	\label{double N: fit}
\end{figure}

\textbf{Update \(\mathbf{x}_t\) by normal approximation and (\(\mathbf{d}\), \(\mathbf{C}\)) by MH}

By using the normal approximated posterior variance as the proposal variance, the acceptance rates are around 0.45 (Since there are 30 neurons, there are 30 acceptance rates).
The trace plots are in Figure \ref{dc MH: trace}
\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[tracd of latent]{\label{fig:a}\includegraphics[width=0.55\textwidth]{xnorm_dcMH.png}}%
		\subfigure[trace of loading and dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{trace_dcMH.png}}%
	}
	\caption{latent Normal Approx. + loading MH, trace plots}
	\label{dc MH: trace}
\end{figure}

Looks much better, although still not perfect yet. The fitting of overall firing rate, latents and dynamics are in Figure \ref{dc MH: fit}.

\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[overall mean firing rate]{\label{fig:a}\includegraphics[width=0.55\textwidth]{FR_dcMH.png}}%
		\subfigure[latents]{\label{fig:b}\includegraphics[width=0.55\textwidth]{latent_dcMH.png}}%
		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{A_dcMH.png}}%
	}
	\caption{latent Normal Approx. + loading MH, fit}
	\label{dc MH: fit}
\end{figure}

See, nearly perfect (especially) for latent.

\textbf{Update both \(\mathbf{x}_t\) and (\(\mathbf{d}\), \(\mathbf{C}\)) by MH}

The acceptance rate (by using the optimal scalar) for latent is 0.6227, which is a bit high (but acceptable). Again, the acceptance rates for loading are around 0.45.

The trace plots are in Figure \ref{double MH: trace}
\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[tracd of latent]{\label{fig:a}\includegraphics[width=0.55\textwidth]{xnorm_doubleMH.png}}%
		\subfigure[trace of loading and dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{trace_doubleMH.png}}%
	}
	\caption{both MH, trace plots}
	\label{double MH: trace}
\end{figure}

The trace for dynamics is terrible... Maybe this is caused by high dimension of latent (\(d=6000\) in this case).

The fitting of overall firing rate, latents and dynamics are in Figure \ref{double MH: fit}.

\begin{figure}[h!]
	\makebox[\linewidth][c]{%
		\centering
		\subfigure[overall mean firing rate]{\label{fig:a}\includegraphics[width=0.55\textwidth]{FR_doubleMH.png}}%
		\subfigure[latents]{\label{fig:b}\includegraphics[width=0.55\textwidth]{latent_doubleMH.png}}%
		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{A_doubleMH.png}}%
	}
	\caption{both MH, fit}
	\label{double MH: fit}
\end{figure}

OK, the fitting sucks...

One natural idea is to combine MH and normal approximation in latent: at first few steps, use normal approximation at conditional posterior mode to bring things into somewhere close to mode quickly. Then use MH to stabilize everything.

>>>>

In summary, Let
\(\mathbf{\Theta} = \left\{ \mathbf{z},\ \mathbf{d},\mathbf{C},\left\{ \mathbf{x}_{t} \right\}_{t = 1}^{T},\mathbf{x}_{0},\mathbf{Q}_{0},\mathbf{A},\ \mathbf{b},\mathbf{Q}\  \right\}\)
be the set of parameters for LDS. The number of observation/ neuron is
\(N\), and they can group into \(J\) clusters. In each cluster, there
are \(p\) latent state vectors. The recording length is \(T\).
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	\(\mathbf{z} \in \mathbb{Z}^{N}\): cluster index for each neuron/ observation,
	with \(z_{i} \in \{ 1,\ldots,J\}\) for \(i = 1,\ldots,N\).
	\item
	\(\mathbf{d} \in \mathbb{R}^{N}\) and \(\mathbf{C} \in \mathbb{R}^{N \times p}\): baseline
	\& loading of the latent
	\item
	\(\mathbf{x}_{t} \in \mathbb{R}^{Jp}\): all latent at \(t\).
	\item
	\(\mathbf{x}_{t}^{(j)} \in \mathbb{R}^{p}\) means the latent in cluster \(j\) at
	time \(t\).
	\item
	\(\mathbf{x}_{0} \in \mathbb{R}^{Jp}\) and
	\(\mathbf{Q}_{0} \in \mathbb{R}^{Jp \times Jp}\): mean and covariance of
	\(\mathbf{x}_{1}\). For simplicity, assume \(\mathbf{Q}_{0}\) is known.
	\item
	\(\mathbf{A} \in \mathbb{R}^{Jp \times Jp},\ \mathbf{b} \in \mathbb{R}^{Jp},\mathbf{Q} \in \mathbb{R}^{Jp \times Jp}\):
	linear dynamics of the latent.
\end{enumerate}

@@

The conditional priors for these parameters:
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	Latent vectors \(\left\{ \mathbf{x}_{t} \right\}_{t=1}^T\): the conditional prior is
	defined by
	\[\mathbf{x}_{1} \sim N(\mathbf{x}_{0},\mathbf{Q}_{0})\]
	\[\mathbf{x}_{t + 1}|\mathbf{x}_{t} \sim N(\mathbf{A}\mathbf{x}_{t} + \mathbf{b},\mathbf{Q})\]
	\item
	Initials \(\mathbf{x}_{0}\): assume there are \(J\) clusters,
	\[\mathbf{x}_{0} \sim N(\bm{\mu}_{\mathbf{x}_{00}},\ \mathbf{\Sigma}_{\mathbf{x}_{00}})\]
	, where \(\bm{\mu}_{\mathbf{x}_{00}} = \mathbf{0}_{Jp}\) and \(\mathbf{\Sigma}_{\mathbf{x}_{00}} = \mathbf{I}_{Jp}\)
	\item
	Linear mapping (loading) for latent vectors
	\(\left\{ d_{i} \right\}_{i = 1}^{N}\) and
	\(\left\{ \mathbf{c}_{i} \right\}_{i = 1}^{N}\):
	\[\left( d_{i},\mathbf{c}'_{i} \right)' \sim N(\bm{\mu}_{\text{dc}}^{\left( z_{i} \right)},\mathbf{\Sigma}_{\text{dc}}^{(z_{i})})\]
	\item
	Mean and covariance for loading in each cluster
	Mean and covariance for loading in each cluster
	\(\left\{ \bm{\mu}_{\text{dc}}^{(j)} \right\}_{j}\) and
	\(\left\{ \mathbf{\Sigma}_{\text{dc}}^{(j)} \right\}_{j}\):
	\[\bm{\mu}_{\text{dc}}^{(j)} \sim N(\mathbf{\delta}_{dc0},\mathbf{T}_{dc0})\]
	, where \(\bm{\delta}_{dc0} = \mathbf{0}_{p + 1}\) and
	\(\mathbf{T}_{dc0} = \mathbf{I}_{p + 1}\)

@@

	, where \(\mathbf{f}_0 = vec(\mathbf{F}_0)\).If the number of cluster is \(J\), \(\nu_{\mathbf{Q}_{0}} = Jp + 2\),
	\(\Psi_{\mathbf{Q}_{0}} = \mathbf{I}_{Jp} \times 10^{- 4}\). (To make the mean of \(\mathbf{Q}\) loosely centered around
	\(\mathbf{I}_{Jp} \times 10^{- 4}\)), \(\mathbf{F}_0 = \begin{pmatrix}
		\mathbf{0}'_{Jp} \\
		\mathbf{I}_{Jp} \\
	\end{pmatrix}\) and \(\bm{\Lambda}_0 = \mathbf{I}_{Jp + 1}\)

@@

, where
\(\mathbf{E} = \left( \mathbf{\epsilon}_{1},\ldots,\mathbf{\epsilon}_{T - 1} \right)^{'}\) and \(\mathbf{\epsilon} \sim N(0,\mathbf{Q})\). Then posteriors are
\begin{align*}
	\mathbf{Q}|\mathbf{Y}_{\mathbf{bA}},\mathbf{X}_{\mathbf{bA}},\ldots &\sim W^{-1}(\Psi_{\mathbf{Q}}, \nu_{\mathbf{Q}})\\
	\mathbf{f}|\mathbf{Y}_{\mathbf{bA}},\mathbf{X}_{\mathbf{bA}},\mathbf{Q},\ldots &\sim N(vec(\mathbf{F}_\mathbf{bA}),  \mathbf{Q}\bigotimes \bm{\Lambda}_\mathbf{bA}^{-1})
\end{align*}
, with parameters be:
\begin{align*}
	\Psi_{\mathbf{Q}} &= \Psi_{\mathbf{Q}_{0}} + (\mathbf{Y}_\mathbf{bA} - \mathbf{X}_\mathbf{bA}\mathbf{F}_\mathbf{bA})'(\mathbf{Y}_\mathbf{bA} - \mathbf{X}_\mathbf{bA}\mathbf{F}_\mathbf{bA}) + (\mathbf{F}_\mathbf{bA} - \mathbf{F}_0)'\bm{\Lambda}_0(\mathbf{F}_\mathbf{bA} - \mathbf{F}_0)\\
	\nu_{\mathbf{Q}} &= \nu_{\mathbf{Q}_{0}} + T - 1\\
	\mathbf{F}_\mathbf{bA} &= (\mathbf{X}'_\mathbf{bA}\mathbf{X}_\mathbf{bA} + \bm{\Lambda}_0)^{-1}(\mathbf{X}'_\mathbf{bA}\mathbf{Y}_\mathbf{bA} + \bm{\Lambda}_0\mathbf{F}_0)\\
	\bm{\Lambda}_\mathbf{bA} &= \mathbf{X}'_\mathbf{bA}\mathbf{X}_\mathbf{bA} + \bm{\Lambda}_0
\end{align*}

%\section{Model Check [Problems for \(\mathbf{b}\) and \(\mathbf{A}\) partially solved]} \label{problems}
%
%\subsection{Estimation of \(\mathbf{d}\) and \(\mathbf{C}\) alone}
%First, turn off all others except for loading (related) parameters and set them to be true values. There are 2 versions: (1) update priors as in subsection \ref{loading prior} and (2) no update of priors, and just set the prior for each as the standard normal.
%For the estimation with prior update, the results are mean from iteration 50 to 100 (Figure \ref{loading alone}).
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[]{\includegraphics[width=0.6\textwidth]{image016.png}}%
%		\subfigure[]{\includegraphics[width=0.6\textwidth]{image017.png}}%
%	}
%	\caption{estimation of loading}
%	\label{loading alone}
%\end{figure}
%
%The mixing of convergence for loading priors (in subsection \ref{loading prior}) are fine. The update of priors don't influence results a lot, but this will help clustering a lot. It seems estimation of loading is fine. Let's see what happens when estimation of latents \(\mathbf{x}_t\) is on at the same time.
%
%\subsection{Estimation of \(\mathbf{d}\), \(\mathbf{C}\) and \(\mathbf{x}_{t}\)}
%When the estimation of \(\mathbf{x}_t\) is on, we need more iterations. The results are mean from iteration 500 to 1000 (Figure \ref{loading and latent}).
%
%The estimation of loading:
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[]{\includegraphics[width=0.6\textwidth]{image018.png}}%
%		\subfigure[]{\includegraphics[width=0.6\textwidth]{image020.png}}%
%	}
%	\caption{estimation of loading, with latents on}
%	\label{loading and latent}
%\end{figure}
%
%And the estimation of corresponding latents (Figure \ref{loading and latents}):
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[no update of prior]{\includegraphics[width=0.6\textwidth]{image019.png}}%
%		\subfigure[with update of prior]{\includegraphics[width=0.6\textwidth]{image021.png}}%
%	}
%	\caption{estimation of loading}
%	\label{loading and latents}
%\end{figure}
%
%Things are still fine. The remaining parts may influence latent estimations are dynamics (\(\mathbf{b}\), \(\mathbf{A}\)) and prior process noise \(\mathbf{Q}\).
%
%\subsection{Estimation of \(\mathbf{b}\), \(\mathbf{A}\), \(\mathbf{Q}\) and \(\mathbf{x}_t\)}
%I show the results of block-diagonal (Figure \ref{bAQX_blkDiag}) and full (Figure \ref{bAQX_full}) \(\mathbf{Q}\). These results are from 2000 MCMC samples. The latent and dynamics are from average from 500 to 2000 iterations.
%
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[trace]{\label{fig:a}\includegraphics[width=0.55\textwidth]{bAX_blkDiag_trace.png}}%
%		\subfigure[latent]{\label{fig:b}\includegraphics[width=0.55\textwidth]{bAX_blkDiag_latent.png}}%
%		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{bAX_blkDiag_A.png}}%
%	}
%	\caption{sampling of $\mathbf{b}$, $\mathbf{A}$, $\mathbf{Q}$ and $\mathbf{x}_t$, block-diagonal process noise}
%	\label{bAQX_blkDiag}
%\end{figure}
%
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[trace]{\label{fig:a}\includegraphics[width=0.55\textwidth]{bAX_full_trace.png}}%
%		\subfigure[latent]{\label{fig:b}\includegraphics[width=0.55\textwidth]{bAX_full_latent.png}}%
%		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{bAX_full_A.png}}%
%	}
%	\caption{sampling of $\mathbf{b}$, $\mathbf{A}$, $\mathbf{Q}$ and $\mathbf{x}_t$, full process noise}
%	\label{bAQX_full}
%\end{figure}
%
%Both block-diagonal and full looks fine, but the mixing of  full process noise version looks better.

@@

\textbf{\textcolor{red}{All the simulation results are old and problematic. Update them later.}}

%There are two set of simulation examples. The first example is generated from LDS model directly, while in the second example the latents are generated directly without explicit specifying linear dynamics. In all following results, I fit things with both \href{https://github.com/weigcdsb/state-space-clustering/tree/main/LDS/blkDiag}{block-diagonal} and \href{https://github.com/weigcdsb/state-space-clustering/tree/main/LDS/full}{full} \(\mathbf{Q}\). The fitting results for block-diagonal version is a bit better, because the underlying true \(\mathbf{Q}\) is diagonal. In the following part, I only show results from block-diagonal fitting for labeled data. For unlabeled data, i.e. clustering, all the results can be found in this \href{https://github.com/weigcdsb/state-space-clustering/tree/main/results/gif}{folder}.
%
%\subsection{Simulation 1: Generate from LDS Directly}
%In this simulation, there are 3 clusters with 10 neurons in each cluster. The dimension of latents in each cluster is 2. In the linear dynamics, the bias term \(\mathbf{b}\) is zero. There's no within-population interaction but has some weak between-population interactions. In other words, the linear dynamics matrix \(\mathbf{A}\) is roughly diagonal. The details of simulation can be found in the first section of the \href{https://github.com/weigcdsb/state-space-clustering/blob/main/LDS/blkDiag/lds_sample_DP_blkDiag.m}{simulation 1}.
%
%\subsubsection{Labeled Data: No Clustering}
%The code for fitting is \href{https://github.com/weigcdsb/state-space-clustering/blob/main/LDS/blkDiag/lds_sample_DP_blkDiag.m}{here}. 
%
%Here, I first compare 10k iterations for three methods (consider high dimension of \(\mathbf{x}_t\)): (1) both \(\mathbf{x}_t\) and loading (\(\mathbf{d}\) and \(\mathbf{C}\)) are updated by normal approximation. (2) update \(\mathbf{x}_t\) by normal approximation, but update loading by MH. (3) update both \(\mathbf{x}_t\) and loading by MH.
%
%The fitting results are average of iteration 5k to 10k.
%
%\textbf{Update \(\mathbf{x}_t\) and (\(\mathbf{d}\), \(\mathbf{C}\)) by normal approximation}
%
%The trace plots are in Figure \ref{double N: trace}
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[tracd of latent]{\label{fig:a}\includegraphics[width=0.55\textwidth]{xnorm_doubleN.png}}%
%		\subfigure[trace of loading and dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{trace_doubleN.png}}%
%	}
%	\caption{both Normal Approx., trace plots}
%	\label{double N: trace}
%\end{figure}
%
%Well, the chain hasn't achieve the stationary distribution. So there's no need to see further results. But to show the improvement and for comparison, I still show the fitting of overall firing rate, latents and dynamics in Figure \ref{double N: fit}.
%
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[overall mean firing rate]{\label{fig:a}\includegraphics[width=0.55\textwidth]{FR_doubleN.png}}%
%		\subfigure[latents]{\label{fig:b}\includegraphics[width=0.55\textwidth]{latent_doubleN.png}}%
%		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{A_doubleN.png}}%
%	}
%	\caption{both Normal Approx., fit}
%	\label{double N: fit}
%\end{figure}
%
%\textbf{Update \(\mathbf{x}_t\) by normal approximation and (\(\mathbf{d}\), \(\mathbf{C}\)) by MH}
%
%By using the normal approximated posterior variance as the proposal variance, the acceptance rates are around 0.45 (Since there are 30 neurons, there are 30 acceptance rates).
%The trace plots are in Figure \ref{dc MH: trace}
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[tracd of latent]{\label{fig:a}\includegraphics[width=0.55\textwidth]{xnorm_dcMH.png}}%
%		\subfigure[trace of loading and dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{trace_dcMH.png}}%
%	}
%	\caption{latent Normal Approx. + loading MH, trace plots}
%	\label{dc MH: trace}
%\end{figure}
%
%Looks much better, although still not perfect yet. The fitting of overall firing rate, latents and dynamics are in Figure \ref{dc MH: fit}.
%
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[overall mean firing rate]{\label{fig:a}\includegraphics[width=0.55\textwidth]{FR_dcMH.png}}%
%		\subfigure[latents]{\label{fig:b}\includegraphics[width=0.55\textwidth]{latent_dcMH.png}}%
%		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{A_dcMH.png}}%
%	}
%	\caption{latent Normal Approx. + loading MH, fit}
%	\label{dc MH: fit}
%\end{figure}
%
%See, nearly perfect (especially) for latent.
%
%\textbf{Update both \(\mathbf{x}_t\) and (\(\mathbf{d}\), \(\mathbf{C}\)) by MH}
%
%The acceptance rate (by using the optimal scalar) for latent is 0.6227, which is a bit high (but acceptable). Again, the acceptance rates for loading are around 0.45.
%
%The trace plots are in Figure \ref{double MH: trace}
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[tracd of latent]{\label{fig:a}\includegraphics[width=0.55\textwidth]{xnorm_doubleMH.png}}%
%		\subfigure[trace of loading and dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{trace_doubleMH.png}}%
%	}
%	\caption{both MH, trace plots}
%	\label{double MH: trace}
%\end{figure}
%
%The trace for dynamics is terrible... Maybe this is caused by high dimension of latent (\(d=6000\) in this case).
%
%The fitting of overall firing rate, latents and dynamics are in Figure \ref{double MH: fit}.
%
%\begin{figure}[h!]
%	\makebox[\linewidth][c]{%
%		\centering
%		\subfigure[overall mean firing rate]{\label{fig:a}\includegraphics[width=0.55\textwidth]{FR_doubleMH.png}}%
%		\subfigure[latents]{\label{fig:b}\includegraphics[width=0.55\textwidth]{latent_doubleMH.png}}%
%		\subfigure[dynamics \(\mathbf{A}\)]{\label{fig:c}\includegraphics[width=0.55\textwidth]{A_doubleMH.png}}%
%	}
%	\caption{both MH, fit}
